---
title: "Twin Cities Station Mapping"
format: gfm
editor: visual
---

## Research Question

Which demographic groups are impacted by light rail openings in the United States?

## Research Context

We're focused on the "Twin Cities", Minneapolis and St. Paul, Minnesota and the impact of the opening of the METRO Blue and METRO Green lines.

Timeline: 2000-2008

Hypothesis: Light rail openings decrease the pollution around stations

-   Reasoning: While light rail openings gather commuters, only two of the stations have parking, so on average, the net PM2.5 pollution should in theory be lower. In addition, many of these commuters are electing to use the light rail instead of their own transportation, further contributing to a reduction in PM2.5

-   Confounding Factors:

    -   Power Plants

    -   Recycling Centers

    -   Refineries

    -   Factories

    -   Heating Plants

    -   Paper Mills

    -   Meteorological Factors

    -   Local/Federal/State Policies

        -   **EPA's Heavy-Duty Engine and Vehicle Standards (2001)**: These standards, implemented in 2007, mandated reductions in particulate matter emissions from diesel engines.

        -   **Clean Diesel Trucks and Buses Rule (2001)**: This rule aimed to reduce diesel particulate matter emissions by over 90% from 2007 onwards.

        -   **Nonroad Diesel Rule (2004)**: This rule set stringent emission standards for nonroad diesel engines, which took effect in stages starting in 2008.

        -   **Minnesota Mercury Reduction Act (2006)**: While primarily targeting mercury, this act also required reductions in other pollutants, including PM2.5, from power plants.

        -   **Minnesota Clean Car Rule (2007)**: This rule adopted California’s stricter vehicle emissions standards, which included measures to reduce PM2.5 emissions from cars and light trucks.

        -   **Next Generation Energy Act (2007)**: This act set aggressive targets for reducing greenhouse gas emissions and included measures to increase energy efficiency and the use of renewable energy sources, indirectly impacting PM2.5 levels.

        -   **Metropolitan Council's Regional Development Framework (2004)**: This framework promoted sustainable development and reduced sprawl, which can help reduce vehicle emissions and PM2.5 levels.

        -   **Minneapolis Green Fleet Policy (2007)**: This policy aimed to reduce emissions from the city’s fleet of vehicles, including particulate matter.

        -   **Clean Air Minnesota Initiative (2005)**: This public-private partnership worked to reduce air pollution through voluntary measures, including efforts to reduce PM2.5 emissions.

        -   **Minnesota's Renewable Energy Standard (2007)**: Mandated that 25% of the state's electricity come from renewable sources by 2025, reducing reliance on coal-fired power plants, a major source of PM2.5.

        -   **Vehicle Emissions Inspection and Maintenance Programs**: These programs, though primarily aimed at reducing smog-forming pollutants, also contributed to lower PM2.5 levels by ensuring vehicles were properly maintained and operating efficiently.

```{r}
#| warning: false
#| message: false
library("knitr")
figure1 <- read.csv("Sources of Pollution - Twin Cities - Sheet1 (2).csv")
kable(figure1)
```

## Green & Blue Line Stations

Below is a list of all 37 stations for the METRO Green and Blue Lines.

```{r}
figure2 <- read.csv("Twin Cities Stations - Sheet1.csv")
kable(figure2)
```

## Station Information

This code reads the data from csv files we created that contains the names of each station . The METRO Blue and Green lines in total have 37 stations. This file has the name of each station, their addresses, opening dates, and whether or not they have parking.

```{r}
#| warning: false
#| message: false
library("tidyverse")
library("ggmap") 

station_data <- read.csv("Twin Cities Stations - Sheet1.csv") %>%
  mutate(Station2=paste(Stations, "Station Metro Transit, Minnesota"))
```

This next segments use a Google API key in order to get the locations of each station of the light rail in latitude/longitude coordinates, then fixes a few select points where the Google API failed to get a precise location.

```{r}
#| eval: false
register_google(key = "GoogleAPIKey", write = TRUE)

addr.geo <- mutate_geocode(station_data, location = Station2, output = "latlona")
```

```{r}
#| eval: false
geo <- addr.geo %>%
  mutate(lat2 = ifelse(address=="bloomington, mn, usa", 44.85639, lat)) %>%
  mutate(lon2 = ifelse(address=="bloomington, mn, usa", -93.22628, lon)) %>%
  mutate(lat2 = ifelse(address=="minnesota, usa", 44.95648, lat2)) %>%
  mutate(lon2 = ifelse(address=="minnesota, usa", -93.17874, lon2)) %>%
  mutate(lat2 = ifelse(Station2=="Western Avenue station Metro Transit, Minnesota", 44.95586, lat2)) %>%
  mutate(lon2 = ifelse(Station2=="Western Avenue station Metro Transit, Minnesota", -93.11708, lon2))

```

## Plotting the Stations

This next code segment plots each of the station locations (in latitude/longitude) we extracted in the previous stage

```{r}
#| warning: false
#| message: false
library("terra")
geo <- read.csv("stations_with_locations.csv")

sample_coords <- cbind(geo$lon2, geo$lat2)
lr_stations <- vect(sample_coords)
crdref <- "+proj=longlat +datum=WGS84"
pts <- vect(sample_coords, crs=crdref)
plot(pts)

```

## Plotting the Stations (Part 2)

This code chunk displays the station locations on top of a real map with a 500m buffer circle around each station location. These buffers and their sizes are to help to later extract the average PM2.5 and control data for each of the stations individually while limiting the influence of potential other sources of PM2.5 as well as overlap with other stations. In addition, these buffers are necessary to get an average reading of all the data near the station rather than relying on a single point, which could lead to outliers.

```{r}
library("terra")
library("maptiles")
geo <- read.csv("stations_with_locations.csv")

pts_buffer <- buffer(pts, width = 500) # Width is measured in meters
plot(pts_buffer)

#writeVector(pts_buffer, "Station_Buffers.shp")

tc_lr <- vect("lr_1km_buff/lr_1km_buff.shp")
tc_lr_line <- aggregate(tc_lr, dissolve = TRUE)
lr_project <- project(tc_lr_line, "+proj=longlat + ellps = WGS84 +datum = WGS84 + nodefs")

lrc <- centroids(lr_project, inside = FALSE)

pts_buffer1 <- buffer(lrc, width = 10000)
extent <- buffer(pts, width = 600)

bg <- get_tiles(ext(extent))
plot(bg)
points(pts)
lines(pts_buffer1, col = "blue")
lines(pts_buffer, col = "red")
```

## Find Sources of Pollution near Light Rail Routes and its Addresses

This file has the name of each factory, their addresses, opening dates, and whether or not they have parking.

```{r}
library("tidyverse")
library("ggmap") 

factory_data <- read.csv("Sources of Pollution - Twin Cities - Sheet1 (2).csv") %>%
  mutate(Factory2=paste(Factories, "Factory, Minnesota"))
```

This next segments use a Google API key in order to get the locations of each station of the light rail in latitude/longitude coordinates, then fixes a few select points where the Google API failed to get a precise location.

```{r}
#| eval: false
register_google(key = "GoogleAPIKey", write = TRUE)

addr.geo <- mutate_geocode(factory_data, location = Factory2, output = "latlona")
```

```{r}
#| eval: false
geo2 <- addr.geo %>%
  mutate(lat2 = ifelse(Factories=="Flint Hills Resources Line Bend Refinery, Rosemount, MN", 44.76432, lat)) %>%
  mutate(lon2 = ifelse(Factories=="Flint Hills Resources Line Bend Refinery, Rosemount, MN", -93.03947, lon)) %>%
  mutate(lat2 = ifelse(Factories=="Koch Industries (Flint Hills Resources), Rosemount, MN", 44.76424, lat2)) %>%
  mutate(lon2 = ifelse(Factories=="Koch Industries (Flint Hills Resources), Rosemount, MN", -93.03943, lon2))

write.csv(geo2, "pollution_locations.csv")
```

## Plotting the Factory Sources of Pollution

This code plots nearby sources of pollution (in purple), and where they are located in comparison to the light rail itself. This will be factored in later when deciding which stations may be outliers in terms of PM2.5 air pollution.

```{r}
library("terra")
geo <- read.csv("stations_with_locations.csv")
geo2 <- read.csv("pollution_locations.csv")

sample_coords <- cbind(geo$lon2, geo$lat2)
lr_stations <- vect(sample_coords)
crdref <- "+proj=longlat +datum=WGS84"
pts <- vect(sample_coords, crs=crdref)

pol_coords <- cbind(geo2$lon2, geo2$lat2)
crdref <- "+proj=longlat +datum=WGS84"
pol_pts <- vect(pol_coords, crs=crdref)

library("maptiles")

pts_buffer <- buffer(pts, width = 500) # Width is measured in meters

lrc <- centroids(lr_project, inside = FALSE)

pts_buffer1 <- buffer(lrc, width = 10000)
extent <- buffer(pts, width = 600)

bg <- get_tiles(ext(extent))
plot(bg)
points(pts)
lines(pts_buffer, col = "blue")
points(pol_pts, col = "purple", cex = 1.5)
```

## Extracting Meteorology Data

This code extracts all of the meteorology data provided by GLDAS for all of the 3,288 days in our period of interest, cropped to the areas within each of the station buffers. It then takes the averages of each meteorological factor in each station then writes them to individual dataframes to be combined in the next section. We extracted all the meteorological factors for convenience, but we were mainly interested in the temperature, humidity, and wind speed at each station, as these are the meteorological factors that would have the most impact on PM2.5 pollution.

```{r}
#| eval: false

library("terra")
library("tidyverse")

#makes of list of files in that folder
files<-dir("G:/Shared drives/2024 FIRE Light Rail/DATA/GLDAS/")

for(i in 1:3288){
  r<-rast(paste0("G:/Shared drives/2024 FIRE Light Rail/DATA/GLDAS/", files[i]))

  names(r)
  #variables in page 19 of manual
  #https://hydro1.gesdisc.eosdis.nasa.gov/data/GLDAS/GLDAS_CLSM025_D.2.0/doc       /README_GLDAS2.pdf
  #Snowf_tavg<-r[[6]]
  #plot(Snowf_tavg)

  station_buffers<-vect("Station_Buffers.shp")

  #crops raster to contain only buffers around stations
  int<-crop(r, station_buffers,
            snap="in",
            mask=TRUE)

  #convert cropped raster into dataframe and find average value
  metdf<-terra::extract(int, sta, fun="mean", na.rm=TRUE)  %>% 
    summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE))) %>%
    select(-ID)

  metdf$date<-files[i]
  
  write.csv(metdf, paste0("TC_Meteorology_Data/", files[i],".csv"), row.names = F)

}
```

## Combining the data

This code combines all of the previous data we've gathered into a single dataframe, along with adding in some additional variables. This dataframe includes each station name, each station id, the date, PM2.5, meteorological variables, the month, day of the week, and whether or not the day is a holiday.

```{r}
#| warning: false
#| message: false
#| eval: false

library("dplyr")
library("data.table")
path<-"C:/Users/rygel/Documents/team-twin-cities/TC_Meteorology_Data/"
days<-dir(path) #makes a vector of folder names

setwd("C:/Users/rygel/Documents/team-twin-cities/TC_Meteorology_Data/")

combined_files <- bind_rows(lapply(days, fread))

setwd("C:/Users/rygel/Documents/team-twin-cities/")

write.csv(combined_files, "Full_Meteorology_Data.csv")

stations <- read.csv("Station Names and IDs.csv")

path<-"C:/Users/rygel/Documents/team-twin-cities/Twin_Cities_PM25/"
months<-dir(path) #makes a vector of folder names

setwd("C:/Users/rygel/Documents/team-twin-cities/Twin_Cities_PM25/")


combined_files <- bind_rows(lapply(months, fread))

write.csv(combined_files, "Full_PM25_Data.csv", row.names = F)

setwd("C:/Users/rygel/Documents/team-twin-cities/")

PM25 <- read.csv("Full_PM25_Data.csv")
Holidays <- read.csv("major_holidays_2000_2025.csv")
Holidays = subset(Holidays, select = -year)
Weather <- read.csv("Full_Meteorology_Data.csv")
Weather = subset(Weather, select = -X)
Weather = subset(Weather, select = -V1)
stations = read.csv("Station Names and IDs.csv")

weather_fixed_date <- Weather %>%
  mutate(date = stringr::str_extract(date, "[0-9]{4}[0-9]{2}[0-9]{2}")) %>%
  mutate(date = paste(substr(date, 1, 4), "-", substr(date, 5, 6), "-", substr(date, 7, nchar(date)), sep = ""))

pm25_fixed_date <- PM25 %>%
  mutate(date = stringr::str_extract(date, "[0-9]{4}[0-9]{2}[0-9]{2}")) %>%
  mutate(date = paste(substr(date, 1, 4), "-", substr(date, 5, 6), "-", substr(date, 7, nchar(date)), sep = ""))

pm25_stations = merge(pm25_fixed_date, stations, by = "station_num")

add_weather = merge(pm25_stations, weather_fixed_date, by = "date", all.x = TRUE, all.y = FALSE)

add_holidays = merge(add_weather, Holidays, by = "date", all.x = TRUE, all.y = FALSE)

full_fixed_cols <- add_holidays %>%
  mutate(holiday = ifelse(is.na(holiday), FALSE, TRUE)) %>%
  mutate(day_of_week = weekdays(as.Date(date))) %>%
  mutate(month = months(as.Date(date)))

date_stations_sorted <- full_fixed_cols[order(full_fixed_cols$date, full_fixed_cols$station_num),]

write.csv(date_stations_sorted, "Mega_Dataframe.csv", row.names = F)
```

## Plotting PM2.5 of the different stations over time

This code plots a boxplot for the PM2.5 on each date from 2000-2008 for each station as well as finds the differences in average PM2.5 before and after the opening of the METRO Blue Line on June 14th, 2004. Using this data as well as factoring in parking and other pollution sources, we have decided to remove stations 1, 16, 28, and 34 from our data.

```{r}
#| warning: false
library(ggplot2)

all_data = read.csv("Mega_Dataframe.csv")
half_data = head(all_data, 120398)
truncated_data = head(all_data, 60199)
truncated_data2 = head(tail(all_data, -60199), 60199)

ggplot(data = half_data, aes(x = station_num, y = pm25)) +
  geom_boxplot(outlier.shape = NA, aes(group = station_num, color = factor(station_num))) +
  ylim(0, 25)

#ggplot(data = truncated_data, aes(x = station_num, y = pm25)) +
  #geom_boxplot(outlier.shape = NA, aes(group = station_num, color = #factor(station_num))) +
  #ylim(0, 25)

#ggplot(data = truncated_data2, aes(x = station_num, y = pm25)) +
  #geom_boxplot(outlier.shape = NA, aes(group = station_num, color = #factor(station_num))) +
  #ylim(0, 25)

before_mean<- aggregate(x=truncated_data$pm25,
                      # Specify group indicator
                      by = list(truncated_data$station_num),      
                      # Specify function (i.e. mean)
                      FUN = mean)

after_mean<- aggregate(x=truncated_data2$pm25,
                      # Specify group indicator
                      by = list(truncated_data2$station_num),      
                      # Specify function (i.e. mean)
                      FUN = mean)


b = before_mean$x
a = after_mean$x

```

## Setting Relevant Dates

The following code sets a list of dates for the relevant start and end dates for the period of interest (June 2000 - June 2008), the opening of the Metro Blue Line, the date when construction on the Blue Line began, and a list of dates when relevant policies that might impact the level of PM2.5 were enacted.

These policies include the Heavy-Duty Engine and Vehicle Standards, which mandated reductions in particulate matter emissions from diesel engines (EPA's Heavy-Duty Engine and Vehicle Standards, 2001), the Nonroad Diesel Rule, which set stringent emission standards for nonroad diesel engines (Nonroad Diesel Rule, 2004), the Minnesota Mercury Reduction Act, which aimed to reduce mercury and other pollutants from power plants (Minnesota Mercury Reduction Act, 2006), the Next Gen Energy Act, which set aggressive targets for reducing greenhouse gas emissions and included measures to increase energy efficiency and the use of renewable energy sources (Next Generation Energy Act, 2007), the 2030 Regional Development Framework, which promoted sustainable development and reduced sprawl (2030 Regional Development Framework, 2004), the Clean Air Minnesota Initiative, which worked to reduce air pollution through voluntary measures (Clean Air Minnesota Initiative, 2005), the Minnesota Renewable Energy Standard, which mandated that 25% of the state's electricity must come from renewable sources by 2025 (Minnesota Renewable Energy Standard, 2007), and the vehicle emissions inspection and maintenance programs, which were aimed at reducing smog-forming pollutants and ensuring vehicles were properly maintained and efficient (Vehicle Emissions Inspection and Maintenance Programs, 2001).

```{r}
#| warning: false
#| message: false
library("tidyverse")
```

```{r}
df <- read.csv("Mega_Dataframe.csv")
```

```{r}
df2 <- df %>%
  mutate(date = as.Date(date, format = '%Y-%m-%d'))
```

```{r}
# Period of Analysis
startdate <- as.Date("2000-06-01", format = '%Y-%m-%d')
enddate <- as.Date("2008-06-01", format = '%Y-%m-%d')
opendate <- as.Date("2004-06-14", format = "%Y-%m-%d")
conststart <- as.Date("2001-01-17", format = "%Y-%m-%d")
#Heavy-Duty Engine and Vehicle Standards
HD_Engine <- as.Date("2007-01-01", format = "%Y-%m-%d")
#Nonroad Diesel Rule
NR_Diesel <- as.Date("2004-06-29", format = "%Y-%m-%d")
#Minnesota Mercury Reduction Act
MC_Reduction <- as.Date("2006-05-11", format = "%Y-%m-%d")
#Next Gen Energy Act
NG_Energy <- as.Date("2007-05-25", format = "%Y-%m-%d")
#2030 Regional Development Framework
RDF <- as.Date("2004-01-01", format = "%Y-%m-%d")
#Clean Air Minnesota Initiative
CA_Init <- as.Date("2003-01-01", format = "%Y-%m-%d")
#Minnesota Renewable Energy Standard
RE_Standard <- as.Date("2001-01-01", format = "%Y-%m-%d")
#Vehicle Emissions Inspection and Maintenance Programs
VE_Programs <- as.Date("2001-04-05", format = "%Y-%m-%d")

```

## Adding DB-OLS Variables

The following code limits the data to the aforementioned period of interest and adds dummy variable columns for the day of the week fixed effects, the opening date of the metro, the start of construction date, the temperature, humidity, and wind speed and their lag polynomials, the time in days and its polynomials, and finally a column for each of the relevant policies from the previous section. The time and weather polynomials are to replicate the methodology of (Chen & Whalley, 2012).

```{r}
df3 <- df2 %>%
  filter(date>=startdate & date <= enddate) %>%
  mutate(MetroOpen = ifelse(date >= opendate, 1, 0)) %>%
  mutate(dow = wday(date)) %>%
  mutate(construction = ifelse(date > conststart & date < opendate, 1, 0)) %>%
  group_by(station_num) %>%
  arrange(station_num, date) %>%
  mutate(t = as.numeric(date-startdate)) %>%
  mutate(t2 = t^2, t3 = t^3, t4 = t^4) %>%
  mutate(l_tair = lag(Tair_f_tavg)) %>%
  mutate(l_tair_2 = l_tair^2) %>%
  mutate(l_tair_3 = l_tair^3) %>%
  mutate(l_tair_4 = l_tair^4) %>%
  mutate(l_qair = lag(Qair_f_tavg)) %>%
  mutate(l_qair_2 = l_qair^2) %>%
  mutate(l_qair_3 = l_qair^3) %>%
  mutate(l_qair_4 = l_qair^4) %>%
  mutate(l_wind = lag(Wind_f_tavg)) %>%
  mutate(l_wind_2 = l_wind^2) %>%
  mutate(l_wind_3 = l_wind^3) %>%
  mutate(l_wind_4 = l_wind^4) %>%
  mutate(CA_Init = ifelse(date >= CA_Init, 1, 0)) %>%
  mutate(HD_Engine = ifelse(date >= HD_Engine, 1, 0)) %>%
  mutate(MC_Reduction = ifelse(date >= MC_Reduction, 1, 0)) %>%
  mutate(NG_Energy = ifelse(date >= NG_Energy, 1, 0)) %>%
  mutate(NR_Diesel = ifelse(date >= NR_Diesel, 1, 0)) %>%
  mutate(RDF = ifelse(date >= RDF, 1, 0)) %>%
  mutate(RE_Standard = ifelse(date >= RE_Standard, 1, 0)) %>%
  mutate(VE_Programs = ifelse(date >= VE_Programs, 1, 0))
```

## Running Regression Models

The following code blocks run various regression models including different control factors each time. The results from these regressions have been collected into the table below:

```{r}
#| warning: false
#| message: false
#| echo: false
library("knitr")
figure1 <- read.csv("Regression Model Results.csv")
kable(figure1)
```

This code block runs a simple regression without any of the control factors (Column "log.PM2.5" in table).

```{r}
#| eval: false
summary(m1 <- lm(log(pm25) ~ MetroOpen, data = df3))
```

This code block runs a regression with the construction, month fixed effects, and day fixed effects added (Column "X.1" in table).

```{r}
#| eval: false
summary(m1 <- lm(log(pm25) ~ MetroOpen +
                   construction +
                   as.factor(month) +
                   as.factor(dow)
                   , data = df3))
```

This regression adds the weather polynomials to the previous one (Column "X.2" in table).

```{r}
#| eval: false
summary(m1 <- lm(log(pm25) ~ MetroOpen +
                   construction +
                   Tair_f_tavg +
                   l_tair +
                   l_tair_2 +
                   l_tair_3 +
                   l_tair_4 +
                   Qair_f_tavg +
                   l_qair +
                   l_qair_2 +
                   l_qair_3 +
                   l_qair_4 +
                   Wind_f_tavg +
                   l_wind +
                   l_wind_2 +
                   l_wind_3 +
                   l_wind_4 +
                   as.factor(month) +
                   as.factor(dow)
                   , data = df3))
```

This block adds the holiday control variable into the regression (Column "X.3" in table).

```{r}
#| eval: false
summary(m1 <- lm(log(pm25) ~ MetroOpen +
                   construction +
                   Tair_f_tavg +
                   l_tair +
                   l_tair_2 +
                   l_tair_3 +
                   l_tair_4 +
                   Qair_f_tavg +
                   l_qair +
                   l_qair_2 +
                   l_qair_3 +
                   l_qair_4 +
                   Wind_f_tavg +
                   l_wind +
                   l_wind_2 +
                   l_wind_3 +
                   l_wind_4 +
                   holiday +
                   as.factor(month) +
                   as.factor(dow)
                   , data = df3))

```

This block adds the time polynomials into the regression (Column "X.4" in table).

```{r}
#| eval: false
summary(m1 <- lm(log(pm25) ~ MetroOpen +
                   construction +
                   Tair_f_tavg +
                   l_tair +
                   l_tair_2 +
                   l_tair_3 +
                   l_tair_4 +
                   Qair_f_tavg +
                   l_qair +
                   l_qair_2 +
                   l_qair_3 +
                   l_qair_4 +
                   Wind_f_tavg +
                   l_wind +
                   l_wind_2 +
                   l_wind_3 +
                   l_wind_4 +
                   holiday +
                   t +
                   t2 +
                   t3 +
                   t4 +
                   as.factor(month) +
                   as.factor(dow)
                   , data = df3))
```

This block adds the relevant policies for PM2.5 into the regression (Column "X.5" in table).

```{r}
#| eval: false
summary(m1 <- lm(log(pm25) ~ MetroOpen +
                   construction +
                   Tair_f_tavg +
                   l_tair +
                   l_tair_2 +
                   l_tair_3 +
                   l_tair_4 +
                   Qair_f_tavg +
                   l_qair +
                   l_qair_2 +
                   l_qair_3 +
                   l_qair_4 +
                   Wind_f_tavg +
                   l_wind +
                   l_wind_2 +
                   l_wind_3 +
                   l_wind_4 +
                   holiday +
                   t +
                   t2 +
                   t3 +
                   t4 +
                   CA_Init +
                   HD_Engine +
                   MC_Reduction +
                   NG_Energy +
                   NR_Diesel +
                   RDF +
                   RE_Standard +
                   VE_Programs +
                   as.factor(month) +
                   as.factor(dow)
                   , data = df3))
```

Finally, this block removes the stations which were determined to be outliers (1, 16, 28, 34) into the regression (Column "X.6" in table).

```{r}
#| eval: false
df4 <- df3 %>%
  filter(station_num != 1 & station_num != 16 & station_num != 28 & station_num != 34)

summary(m1 <- lm(log(pm25) ~ MetroOpen +
                   construction +
                   Tair_f_tavg +
                   l_tair +
                   l_tair_2 +
                   l_tair_3 +
                   l_tair_4 +
                   Qair_f_tavg +
                   l_qair +
                   l_qair_2 +
                   l_qair_3 +
                   l_qair_4 +
                   Wind_f_tavg +
                   l_wind +
                   l_wind_2 +
                   l_wind_3 +
                   l_wind_4 +
                   holiday +
                   t +
                   t2 +
                   t3 +
                   t4 +
                   CA_Init +
                   HD_Engine +
                   MC_Reduction +
                   NG_Energy +
                   NR_Diesel +
                   RDF +
                   RE_Standard +
                   VE_Programs +
                   as.factor(month) +
                   as.factor(dow)
                   , data = df4))
```

## Timeline

This graph shows the average PM2.5 across all 37 stations over time along with relevant events plotted as vertical lines.

```{r}
df5 <- df3 %>%
  group_by(date) %>%
  summarise(pm25 = mean(pm25))
ggplot(data = df5, aes(x = date, y = pm25)) +
  geom_line(col = "lightblue") +
  geom_vline(aes(xintercept = opendate, col = "Opening Date")) + 
  geom_vline(aes(xintercept = conststart, col = "Construction Date")) + 
  geom_vline(aes(xintercept = CA_Init, col = "CA")) +
  geom_vline(aes(xintercept = HD_Engine, col = "HD")) +
  geom_vline(aes(xintercept = MC_Reduction, col = "MC")) +
  geom_vline(aes(xintercept = NG_Energy, col = "NG"))
```

# Calculating Station-Level Pollution Change

```{r}
library("tidyverse")
library("knitr")
```

This chunk uploads the full data frame required to estimate the impact of light rails on PM2.5 levels surrounding all stations

```{r}
df<-read.csv("Mega_Dataframe.csv") 

df2<-df %>%
  mutate(date=as.Date(date, format='%Y-%m-%d'))

#period of analysis
startdate<-as.Date("2000-06-01", format='%Y-%m-%d')
enddate<-as.Date("2008-06-01", format='%Y-%m-%d')

#opening data of metro
opendate<-as.Date("2004-06-14", format='%Y-%m-%d')

#date when groundbreak starts
conststart <- as.Date("2001-01-17", format = "%Y-%m-%d")

#Heavy-Duty Engine and Vehicle Standards
HD_Engine <- as.Date("2007-01-01", format = "%Y-%m-%d")
#Nonroad Diesel Rule
NR_Diesel <- as.Date("2004-06-29", format = "%Y-%m-%d")
#Minnesota Mercury Reduction Act
MC_Reduction <- as.Date("2006-05-11", format = "%Y-%m-%d")
#Next Gen Energy Act
NG_Energy <- as.Date("2007-05-25", format = "%Y-%m-%d")
#2030 Regional Development Framework
RDF <- as.Date("2004-01-01", format = "%Y-%m-%d")
#Clean Air Minnesota Initiative
CA_Init <- as.Date("2003-01-01", format = "%Y-%m-%d")
#Minnesota Renewable Energy Standard
RE_Standard <- as.Date("2001-01-01", format = "%Y-%m-%d")
#Vehicle Emissions Inspection and Maintenance Programs
VE_Programs <- as.Date("2001-04-05", format = "%Y-%m-%d")

df3 <- df2 %>%
  filter(date>=startdate & date <= enddate) %>%
  mutate(MetroOpen = ifelse(date >= opendate, 1, 0)) %>%
  mutate(dow = wday(date)) %>%
  mutate(construction = ifelse(date > conststart & date < opendate, 1, 0)) %>%
  group_by(station_num) %>%
  arrange(station_num, date) %>%
  mutate(t = as.numeric(date-startdate)) %>%
  mutate(t2 = t^2, t3 = t^3, t4 = t^4) %>%
  mutate(l_tair = lag(Tair_f_tavg)) %>%
  mutate(l_tair_2 = l_tair^2) %>%
  mutate(l_tair_3 = l_tair^3) %>%
  mutate(l_tair_4 = l_tair^4) %>%
  mutate(l_qair = lag(Qair_f_tavg)) %>%
  mutate(l_qair_2 = l_qair^2) %>%
  mutate(l_qair_3 = l_qair^3) %>%
  mutate(l_qair_4 = l_qair^4) %>%
  mutate(l_wind = lag(Wind_f_tavg)) %>%
  mutate(l_wind_2 = l_wind^2) %>%
  mutate(l_wind_3 = l_wind^3) %>%
  mutate(l_wind_4 = l_wind^4) %>%
  mutate(CA_Init = ifelse(date >= CA_Init, 1, 0)) %>%
  mutate(HD_Engine = ifelse(date >= HD_Engine, 1, 0)) %>%
  mutate(MC_Reduction = ifelse(date >= MC_Reduction, 1, 0)) %>%
  mutate(NG_Energy = ifelse(date >= NG_Energy, 1, 0)) %>%
  mutate(NR_Diesel = ifelse(date >= NR_Diesel, 1, 0)) %>%
  mutate(RDF = ifelse(date >= RDF, 1, 0)) %>%
  mutate(RE_Standard = ifelse(date >= RE_Standard, 1, 0)) %>%
  mutate(VE_Programs = ifelse(date >= VE_Programs, 1, 0))
```

This runs a station-level regression using the same regression model as used for column "X.4" in the regression table above.

```{r}
summary(m1 <- lm(log(pm25) ~ MetroOpen:as.factor(station_num) +
                   construction +
                   Tair_f_tavg +
                   l_tair +
                   l_tair_2 +
                   l_tair_3 +
                   l_tair_4 +
                   Qair_f_tavg +
                   l_qair +
                   l_qair_2 +
                   l_qair_3 +
                   l_qair_4 +
                   Wind_f_tavg +
                   l_wind +
                   l_wind_2 +
                   l_wind_3 +
                   l_wind_4 +
                   holiday +
                   t +
                   t2 +
                   t3 +
                   t4 +
                   as.factor(month) +
                   as.factor(dow)
                   , data = df3))
```

The code below prints out the station-level effect and the p-values

```{r}
c <- coef(m1)
len_coef<-length(coef(m1))

#get coefficients of the station-level effect
coef<-coef(m1)[(len_coef-36): len_coef]

#get p values of the station-level effect (p<0.05 is statistically significant)
pval<-summary(m1)$coefficients[,4][(len_coef-3): len_coef]

kable(cbind(coef, pval), digits=2)
```

# Query demographic data

Helpful resource: <https://walker-data.com/tidycensus/articles/basic-usage.html>

Get census key

```{r}
#| eval: false
library(tidycensus)
api_key <- "Your_Census_API_Key"
census_api_key(api_key, install=TRUE, overwrite = T)
readRenviron("~/.Renviron")
```

Identify variables for querying

```{r}
#| eval: false
vars<-load_variables(year=2000, dataset="sf1", cache = TRUE)

```

Query variables (The codes represent all age groups for male and female and will be renamed in the next step). We chose age groups in particular because we believe there could be a disparity between how much the light rail impacts younger vs. older people. Since younger people such as teenagers and college students without their own means of transportation may have a higher likelihood of using the light rail, we believe they may be more affected.

```{r}
#| eval: false
targetvars<-c("P012002", "P012003", "P012004", "P012005", "P012006", "P012007", "P012008", "P012009", "P012010", "P012011", "P012012", "P012013", "P012014", "P012015", "P012016", "P012017", "P012018", "P012019", "P012020", "P012021", "P012022", "P012023", "P012024", "P012025", "P012026", "P012027", "P012028", "P012029", "P012030", "P012031", "P012032", "P012033", "P012034", "P012035", "P012036", "P012037", "P012038", "P012039", "P012040", "P012041", "P012042", "P012043", "P012044", "P012045", "P012046", "P012047", "P012048", "P012049")

age1<-get_decennial(year = 2000, geography = "block", variables=targetvars, state="MN", county="Hennepin", output="wide")
age2<-get_decennial(year = 2000, geography = "block", variables=targetvars, state="MN", county="Ramsey", output="wide")
age<-rbind(age1,age2)
```

Rename columns

```{r}
#| eval: false
age_ranges<-age %>%
  rename(total_male="P012002", total_male_under_5_years="P012003", total_male_5_to_9_years="P012004", total_male_10_to_14_years="P012005", total_male_15_to_17_years="P012006", total_male_18_and_19_years="P012007", total_male_20_years="P012008", total_male_21_years="P012009", total_male_22_to_24_years="P012010", total_male_25_to_29_years="P012011", total_male_30_to_34_years="P012012", total_male_35_to_39_years="P012013", total_male_40_to_44_years="P012014", total_male_45_to_49_years="P012015", total_male_50_to_54_years="P012016", total_male_55_to_59_years="P012017", total_male_60_and_61_years="P012018", total_male_62_to_64_years="P012019", total_male_65_and_66_years="P012020", total_male_67_to_69_years="P012021", total_male_70_to_74_years="P012022", total_male_75_to_79_years="P012023", total_male_80_to_84_years="P012024", total_male_85_years_and_over="P012025", total_female="P012026", total_female_under_5_years="P012027", total_female_5_to_9_years="P012028", total_female_10_to_14_years="P012029", total_female_15_to_17_years="P012030", total_female_18_and_19_years="P012031", total_female_20_years="P012032", total_female_21_years="P012033", total_female_22_to_24_years="P012034", total_female_25_to_29_years="P012035", total_female_30_to_34_years="P012036", total_female_35_to_39_years="P012037", total_female_40_to_44_years="P012038", total_female_45_to_49_years="P012039", total_female_50_to_54_years="P012040", total_female_55_to_59_years="P012041", total_female_60_and_61_years="P012042", total_female_62_to_64_years="P012043", total_female_65_and_66_years="P012044", total_female_67_to_69_years="P012045", total_female_70_to_74_years="P012046", total_female_75_to_79_years="P012047", total_female_80_to_84_years="P012048", total_female_85_years_and_over="P012049")

```

# Calculate affected demographic group

```{r}
#| eval: false
library("terra")
```

```{r}
#| eval: false
buff<-vect("Station_Buffers.shp")
plot(buff)
```

Download shapefiles from [tigris package](https://github.com/walkerke/tigris)

```{r}
#| eval: false
shape<-tigris::blocks(state="MN", county="Ramsey", class="sp", year=2000)
shape2<-tigris::blocks(state="MN", county="Hennepin", class="sp", year=2000)
shapevect1<-vect(shape)
shapevect2<-vect(shape2)
shapevect<-rbind(shapevect1, shapevect2)
shapedf<-as.data.frame(shapevect)

buff2 <- subset(buff, buff$FID == 19)
plot(buff)
lines(buff2, col = 'purple')

```

```{r}
#| eval: false
blockage<-merge(shapevect, age_ranges, by.x="BLKIDFP00", by.y="GEOID")

blockage$blockarea<-expanse(blockage, unit="m")

blockagedf<-as.data.frame(blockage)

summary(blockagedf)

#100 percent match
```

Intersect blocks with buffers

```{r}
#| eval: false
output<-c()

for (i in 0:36) {
  buff2 <- subset(buff, buff$FID == i)

  int<-crop(blockage, buff2)

  int$intarea<-expanse(int, unit="m")

  intdf<-as.data.frame(int)%>%
    mutate(frac_area=intarea/blockarea) %>%
    mutate(total_male=total_male*frac_area, 
         total_male_under_5_years=total_male_under_5_years*frac_area, 
         total_male_5_to_9_years=total_male_5_to_9_years*frac_area, 
         total_male_10_to_14_years=total_male_10_to_14_years*frac_area, 
         total_male_15_to_17_years=total_male_15_to_17_years*frac_area, 
         total_male_18_and_19_years=total_male_18_and_19_years*frac_area, 
         total_male_20_years=total_male_20_years*frac_area, 
         total_male_21_years=total_male_21_years*frac_area, 
         total_male_22_to_24_years=total_male_22_to_24_years*frac_area, 
         total_male_25_to_29_years=total_male_25_to_29_years*frac_area, 
         total_male_30_to_34_years=total_male_30_to_34_years*frac_area, 
         total_male_35_to_39_years=total_male_35_to_39_years*frac_area, 
         total_male_40_to_44_years=total_male_40_to_44_years*frac_area, 
         total_male_45_to_49_years=total_male_45_to_49_years*frac_area, 
         total_male_50_to_54_years=total_male_50_to_54_years*frac_area, 
         total_male_55_to_59_years=total_male_55_to_59_years*frac_area, 
         total_male_60_and_61_years=total_male_60_and_61_years*frac_area, 
         total_male_62_to_64_years=total_male_62_to_64_years*frac_area, 
         total_male_65_and_66_years=total_male_65_and_66_years*frac_area, 
         total_male_67_to_69_years=total_male_67_to_69_years*frac_area, 
         total_male_70_to_74_years=total_male_70_to_74_years*frac_area, 
         total_male_75_to_79_years=total_male_75_to_79_years*frac_area, 
         total_male_80_to_84_years=total_male_80_to_84_years*frac_area, 
         total_male_85_years_and_over=total_male_85_years_and_over*frac_area,
         total_female=total_female*frac_area, 
         total_female_under_5_years=total_female_under_5_years*frac_area, 
         total_female_5_to_9_years=total_female_5_to_9_years*frac_area, 
         total_female_10_to_14_years=total_female_10_to_14_years*frac_area, 
         total_female_15_to_17_years=total_female_15_to_17_years*frac_area, 
         total_female_18_and_19_years=total_female_18_and_19_years*frac_area, 
         total_female_20_years=total_female_20_years*frac_area, 
         total_female_21_years=total_female_21_years*frac_area, 
         total_female_22_to_24_years=total_female_22_to_24_years*frac_area, 
         total_female_25_to_29_years=total_female_25_to_29_years*frac_area, 
         total_female_30_to_34_years=total_female_30_to_34_years*frac_area, 
         total_female_35_to_39_years=total_female_35_to_39_years*frac_area, 
         total_female_40_to_44_years=total_female_40_to_44_years*frac_area, 
         total_female_45_to_49_years=total_female_45_to_49_years*frac_area, 
         total_female_50_to_54_years=total_female_50_to_54_years*frac_area, 
         total_female_55_to_59_years=total_female_55_to_59_years*frac_area, 
         total_female_60_and_61_years=total_female_60_and_61_years*frac_area, 
         total_female_62_to_64_years=total_female_62_to_64_years*frac_area, 
         total_female_65_and_66_years=total_female_65_and_66_years*frac_area, 
         total_female_67_to_69_years=total_female_67_to_69_years*frac_area, 
         total_female_70_to_74_years=total_female_70_to_74_years*frac_area, 
         total_female_75_to_79_years=total_female_75_to_79_years*frac_area, 
         total_female_80_to_84_years=total_female_80_to_84_years*frac_area, 
         total_female_85_years_and_over=total_female_85_years_and_over*frac_area) %>%
    summarize(total_male=sum(total_male), 
          total_male_under_5_years=sum(total_male_under_5_years), 
          total_male_5_to_9_years=sum(total_male_5_to_9_years), 
          total_male_10_to_14_years=sum(total_male_10_to_14_years), 
          total_male_15_to_17_years=sum(total_male_15_to_17_years), 
          total_male_18_and_19_years=sum(total_male_18_and_19_years), 
          total_male_20_years=sum(total_male_20_years), 
          total_male_21_years=sum(total_male_21_years), 
          total_male_22_to_24_years=sum(total_male_22_to_24_years), 
          total_male_25_to_29_years=sum(total_male_25_to_29_years), 
          total_male_30_to_34_years=sum(total_male_30_to_34_years), 
          total_male_35_to_39_years=sum(total_male_35_to_39_years), 
          total_male_40_to_44_years=sum(total_male_40_to_44_years), 
          total_male_45_to_49_years=sum(total_male_45_to_49_years), 
          total_male_50_to_54_years=sum(total_male_50_to_54_years), 
          total_male_55_to_59_years=sum(total_male_55_to_59_years), 
          total_male_60_and_61_years=sum(total_male_60_and_61_years), 
          total_male_62_to_64_years=sum(total_male_62_to_64_years), 
          total_male_65_and_66_years=sum(total_male_65_and_66_years), 
          total_male_67_to_69_years=sum(total_male_67_to_69_years), 
          total_male_70_to_74_years=sum(total_male_70_to_74_years), 
          total_male_75_to_79_years=sum(total_male_75_to_79_years), 
          total_male_80_to_84_years=sum(total_male_80_to_84_years), 
          total_male_85_years_and_over=sum(total_male_85_years_and_over), 
          total_female=sum(total_female), 
          total_female_under_5_years=sum(total_female_under_5_years), 
          total_female_5_to_9_years=sum(total_female_5_to_9_years), 
          total_female_10_to_14_years=sum(total_female_10_to_14_years), 
          total_female_15_to_17_years=sum(total_female_15_to_17_years), 
          total_female_18_and_19_years=sum(total_female_18_and_19_years), 
          total_female_20_years=sum(total_female_20_years), 
          total_female_21_years=sum(total_female_21_years), 
          total_female_22_to_24_years=sum(total_female_22_to_24_years), 
          total_female_25_to_29_years=sum(total_female_25_to_29_years), 
          total_female_30_to_34_years=sum(total_female_30_to_34_years), 
          total_female_35_to_39_years=sum(total_female_35_to_39_years), 
          total_female_40_to_44_years=sum(total_female_40_to_44_years), 
          total_female_45_to_49_years=sum(total_female_45_to_49_years), 
          total_female_50_to_54_years=sum(total_female_50_to_54_years), 
          total_female_55_to_59_years=sum(total_female_55_to_59_years), 
          total_female_60_and_61_years=sum(total_female_60_and_61_years), 
          total_female_62_to_64_years=sum(total_female_62_to_64_years), 
          total_female_65_and_66_years=sum(total_female_65_and_66_years), 
          total_female_67_to_69_years=sum(total_female_67_to_69_years), 
          total_female_70_to_74_years=sum(total_female_70_to_74_years), 
          total_female_75_to_79_years=sum(total_female_75_to_79_years), 
          total_female_80_to_84_years=sum(total_female_80_to_84_years), 
          total_female_85_years_and_over=sum(total_female_85_years_and_over)) %>%
  mutate(
      pct_total_male = total_male * 100 / total_male,
      pct_total_male_under_5_years = total_male_under_5_years * 100 / total_male,
      pct_total_male_5_to_9_years = total_male_5_to_9_years * 100 / total_male,
      pct_total_male_10_to_14_years = total_male_10_to_14_years * 100 / total_male,
      pct_total_male_15_to_17_years = total_male_15_to_17_years * 100 / total_male,
      pct_total_male_18_and_19_years = total_male_18_and_19_years * 100 / total_male,
      pct_total_male_20_years = total_male_20_years * 100 / total_male,
      pct_total_male_21_years = total_male_21_years * 100 / total_male,
      pct_total_male_22_to_24_years = total_male_22_to_24_years * 100 / total_male,
      pct_total_male_25_to_29_years = total_male_25_to_29_years * 100 / total_male,
      pct_total_male_30_to_34_years = total_male_30_to_34_years * 100 / total_male,
      pct_total_male_35_to_39_years = total_male_35_to_39_years * 100 / total_male,
      pct_total_male_40_to_44_years = total_male_40_to_44_years * 100 / total_male,
      pct_total_male_45_to_49_years = total_male_45_to_49_years * 100 / total_male,
      pct_total_male_50_to_54_years = total_male_50_to_54_years * 100 / total_male,
      pct_total_male_55_to_59_years = total_male_55_to_59_years * 100 / total_male,
      pct_total_male_60_and_61_years = total_male_60_and_61_years * 100 / total_male,
      pct_total_male_62_to_64_years = total_male_62_to_64_years * 100 / total_male,
      pct_total_male_65_and_66_years = total_male_65_and_66_years * 100 / total_male,
      pct_total_male_67_to_69_years = total_male_67_to_69_years * 100 / total_male,
      pct_total_male_70_to_74_years = total_male_70_to_74_years * 100 / total_male,
      pct_total_male_75_to_79_years = total_male_75_to_79_years * 100 / total_male,
      pct_total_male_80_to_84_years = total_male_80_to_84_years * 100 / total_male,
      pct_total_male_85_years_and_over = total_male_85_years_and_over * 100 / total_male,
      pct_total_female = total_female * 100 / total_female,
      pct_total_female_under_5_years = total_female_under_5_years * 100 / total_female,
      pct_total_female_5_to_9_years = total_female_5_to_9_years * 100 / total_female,
      pct_total_female_10_to_14_years = total_female_10_to_14_years * 100 / total_female,
      pct_total_female_15_to_17_years = total_female_15_to_17_years * 100 / total_female,
      pct_total_female_18_and_19_years = total_female_18_and_19_years * 100 / total_female,
      pct_total_female_20_years = total_female_20_years * 100 / total_female,
      pct_total_female_21_years = total_female_21_years * 100 / total_female,
      pct_total_female_22_to_24_years = total_female_22_to_24_years * 100 / total_female,
      pct_total_female_25_to_29_years = total_female_25_to_29_years * 100 / total_female,
      pct_total_female_30_to_34_years = total_female_30_to_34_years * 100 / total_female,
      pct_total_female_35_to_39_years = total_female_35_to_39_years * 100 / total_female,
      pct_total_female_40_to_44_years = total_female_40_to_44_years * 100 / total_female,
      pct_total_female_45_to_49_years = total_female_45_to_49_years * 100 / total_female,
      pct_total_female_50_to_54_years = total_female_50_to_54_years * 100 / total_female,
      pct_total_female_55_to_59_years = total_female_55_to_59_years * 100 / total_female,
      pct_total_female_60_and_61_years = total_female_60_and_61_years * 100 / total_female,
      pct_total_female_62_to_64_years = total_female_62_to_64_years * 100 / total_female,
      pct_total_female_65_and_66_years = total_female_65_and_66_years * 100 / total_female,
      pct_total_female_67_to_69_years = total_female_67_to_69_years * 100 / total_female,
      pct_total_female_70_to_74_years = total_female_70_to_74_years * 100 / total_female,
      pct_total_female_75_to_79_years = total_female_75_to_79_years * 100 / total_female,
      pct_total_female_80_to_84_years = total_female_80_to_84_years * 100 / total_female,
      pct_total_female_85_years_and_over = total_female_85_years_and_over * 100 / total_female
    )
  
  output<-rbind(output, intdf)
  
}
 
write.csv(output, "census_data_full.csv", row.names = F)

```

Summarize demographic groups that live near light rail stations

```{r}
#| eval: false
sum_demog<-output %>%
  select(
  pct_total_male, 
  pct_total_male_under_5_years, 
  pct_total_male_5_to_9_years, 
  pct_total_male_10_to_14_years, 
  pct_total_male_15_to_17_years,
  pct_total_male_18_and_19_years, 
  pct_total_male_20_years, 
  pct_total_male_21_years, 
  pct_total_male_22_to_24_years, 
  pct_total_male_25_to_29_years, 
  pct_total_male_30_to_34_years, 
  pct_total_male_35_to_39_years, 
  pct_total_male_40_to_44_years, 
  pct_total_male_45_to_49_years, 
  pct_total_male_50_to_54_years, 
  pct_total_male_55_to_59_years, 
  pct_total_male_60_and_61_years, 
  pct_total_male_62_to_64_years, 
  pct_total_male_65_and_66_years, 
  pct_total_male_67_to_69_years, 
  pct_total_male_70_to_74_years, 
  pct_total_male_75_to_79_years, 
  pct_total_male_80_to_84_years, 
  pct_total_male_85_years_and_over, 
  pct_total_female, 
  pct_total_female_under_5_years, 
  pct_total_female_5_to_9_years, 
  pct_total_female_10_to_14_years, 
  pct_total_female_15_to_17_years, 
  pct_total_female_18_and_19_years, 
  pct_total_female_20_years, 
  pct_total_female_21_years, 
  pct_total_female_22_to_24_years, 
  pct_total_female_25_to_29_years, 
  pct_total_female_30_to_34_years, 
  pct_total_female_35_to_39_years, 
  pct_total_female_40_to_44_years, 
  pct_total_female_45_to_49_years, 
  pct_total_female_50_to_54_years, 
  pct_total_female_55_to_59_years, 
  pct_total_female_60_and_61_years, 
  pct_total_female_62_to_64_years, 
  pct_total_female_65_and_66_years, 
  pct_total_female_67_to_69_years, 
  pct_total_female_70_to_74_years, 
  pct_total_female_75_to_79_years, 
  pct_total_female_80_to_84_years, 
  pct_total_female_85_years_and_over
)


kable(sum_demog, digits=2)
```

This merges the station-level demographic data with station-level pollution reduction data and

calculates the average emissions that each demographic group sees.

```{r}
#| eval: false
alldf<-cbind(output, coef)
```

```{r}
#| eval: false
reddf<-alldf %>%
  mutate(total_male_change=total_male*coef,
       total_male_under_5_years_change=total_male_under_5_years*coef,
       total_male_5_to_9_years_change=total_male_5_to_9_years*coef,
       total_male_10_to_14_years_change=total_male_10_to_14_years*coef,
       total_male_15_to_17_years_change=total_male_15_to_17_years*coef,
       total_male_18_and_19_years_change=total_male_18_and_19_years*coef,
       total_male_20_years_change=total_male_20_years*coef,
       total_male_21_years_change=total_male_21_years*coef,
       total_male_22_to_24_years_change=total_male_22_to_24_years*coef,
       total_male_25_to_29_years_change=total_male_25_to_29_years*coef,
       total_male_30_to_34_years_change=total_male_30_to_34_years*coef,
       total_male_35_to_39_years_change=total_male_35_to_39_years*coef,
       total_male_40_to_44_years_change=total_male_40_to_44_years*coef,
       total_male_45_to_49_years_change=total_male_45_to_49_years*coef,
       total_male_50_to_54_years_change=total_male_50_to_54_years*coef,
       total_male_55_to_59_years_change=total_male_55_to_59_years*coef,
       total_male_60_and_61_years_change=total_male_60_and_61_years*coef,
       total_male_62_to_64_years_change=total_male_62_to_64_years*coef,
       total_male_65_and_66_years_change=total_male_65_and_66_years*coef,
       total_male_67_to_69_years_change=total_male_67_to_69_years*coef,
       total_male_70_to_74_years_change=total_male_70_to_74_years*coef,
       total_male_75_to_79_years_change=total_male_75_to_79_years*coef,
       total_male_80_to_84_years_change=total_male_80_to_84_years*coef,
       total_male_85_years_and_over_change=total_male_85_years_and_over*coef,
       total_female_change=total_female*coef,
       total_female_under_5_years_change=total_female_under_5_years*coef,
       total_female_5_to_9_years_change=total_female_5_to_9_years*coef,
       total_female_10_to_14_years_change=total_female_10_to_14_years*coef,
       total_female_15_to_17_years_change=total_female_15_to_17_years*coef,
       total_female_18_and_19_years_change=total_female_18_and_19_years*coef,
       total_female_20_years_change=total_female_20_years*coef,
       total_female_21_years_change=total_female_21_years*coef,
       total_female_22_to_24_years_change=total_female_22_to_24_years*coef,
       total_female_25_to_29_years_change=total_female_25_to_29_years*coef,
       total_female_30_to_34_years_change=total_female_30_to_34_years*coef,
       total_female_35_to_39_years_change=total_female_35_to_39_years*coef,
       total_female_40_to_44_years_change=total_female_40_to_44_years*coef,
       total_female_45_to_49_years_change=total_female_45_to_49_years*coef,
       total_female_50_to_54_years_change=total_female_50_to_54_years*coef,
       total_female_55_to_59_years_change=total_female_55_to_59_years*coef,
       total_female_60_and_61_years_change=total_female_60_and_61_years*coef,
       total_female_62_to_64_years_change=total_female_62_to_64_years*coef,
       total_female_65_and_66_years_change=total_female_65_and_66_years*coef,
       total_female_67_to_69_years_change=total_female_67_to_69_years*coef,
       total_female_70_to_74_years_change=total_female_70_to_74_years*coef,
       total_female_75_to_79_years_change=total_female_75_to_79_years*coef,
       total_female_80_to_84_years_change=total_female_80_to_84_years*coef,
       total_female_85_years_and_over_change=total_female_85_years_and_over*coef) %>%
summarize(total_male_change_sum=sum(total_male_change),
          total_male_under_5_years_change_sum=sum(total_male_under_5_years_change),
          total_male_5_to_9_years_change_sum=sum(total_male_5_to_9_years_change),
          total_male_10_to_14_years_change_sum=sum(total_male_10_to_14_years_change),
          total_male_15_to_17_years_change_sum=sum(total_male_15_to_17_years_change),
          total_male_18_and_19_years_change_sum=sum(total_male_18_and_19_years_change),
          total_male_20_years_change_sum=sum(total_male_20_years_change),
          total_male_21_years_change_sum=sum(total_male_21_years_change),
          total_male_22_to_24_years_change_sum=sum(total_male_22_to_24_years_change),
          total_male_25_to_29_years_change_sum=sum(total_male_25_to_29_years_change),
          total_male_30_to_34_years_change_sum=sum(total_male_30_to_34_years_change),
          total_male_35_to_39_years_change_sum=sum(total_male_35_to_39_years_change),
          total_male_40_to_44_years_change_sum=sum(total_male_40_to_44_years_change),
          total_male_45_to_49_years_change_sum=sum(total_male_45_to_49_years_change),
          total_male_50_to_54_years_change_sum=sum(total_male_50_to_54_years_change),
          total_male_55_to_59_years_change_sum=sum(total_male_55_to_59_years_change),
          total_male_60_and_61_years_change_sum=sum(total_male_60_and_61_years_change),
          total_male_62_to_64_years_change_sum=sum(total_male_62_to_64_years_change),
          total_male_65_and_66_years_change_sum=sum(total_male_65_and_66_years_change),
          total_male_67_to_69_years_change_sum=sum(total_male_67_to_69_years_change),
          total_male_70_to_74_years_change_sum=sum(total_male_70_to_74_years_change),
          total_male_75_to_79_years_change_sum=sum(total_male_75_to_79_years_change),
          total_male_80_to_84_years_change_sum=sum(total_male_80_to_84_years_change),
          total_male_85_years_and_over_change_sum=sum(total_male_85_years_and_over_change),
          total_female_change_sum=sum(total_female_change),
          total_female_under_5_years_change_sum=sum(total_female_under_5_years_change),
          total_female_5_to_9_years_change_sum=sum(total_female_5_to_9_years_change),
          total_female_10_to_14_years_change_sum=sum(total_female_10_to_14_years_change),
          total_female_15_to_17_years_change_sum=sum(total_female_15_to_17_years_change),
          total_female_18_and_19_years_change_sum=sum(total_female_18_and_19_years_change),
          total_female_20_years_change_sum=sum(total_female_20_years_change),
          total_female_21_years_change_sum=sum(total_female_21_years_change),
          total_female_22_to_24_years_change_sum=sum(total_female_22_to_24_years_change),
          total_female_25_to_29_years_change_sum=sum(total_female_25_to_29_years_change),
          total_female_30_to_34_years_change_sum=sum(total_female_30_to_34_years_change),
          total_female_35_to_39_years_change_sum=sum(total_female_35_to_39_years_change),
          total_female_40_to_44_years_change_sum=sum(total_female_40_to_44_years_change),
          total_female_45_to_49_years_change_sum=sum(total_female_45_to_49_years_change),
          total_female_50_to_54_years_change_sum=sum(total_female_50_to_54_years_change),
          total_female_55_to_59_years_change_sum=sum(total_female_55_to_59_years_change),
          total_female_60_and_61_years_change_sum=sum(total_female_60_and_61_years_change),
          total_female_62_to_64_years_change_sum=sum(total_female_62_to_64_years_change),
          total_female_65_and_66_years_change_sum=sum(total_female_65_and_66_years_change),
          total_female_67_to_69_years_change_sum=sum(total_female_67_to_69_years_change),
          total_female_70_to_74_years_change_sum=sum(total_female_70_to_74_years_change),
          total_female_75_to_79_years_change_sum=sum(total_female_75_to_79_years_change),
          total_female_80_to_84_years_change_sum=sum(total_female_80_to_84_years_change),
          total_female_85_years_and_over_change_sum=sum(total_female_85_years_and_over_change),
          total_male=sum(total_male),
          total_male_under_5_years=sum(total_male_under_5_years),
          total_male_5_to_9_years=sum(total_male_5_to_9_years),
          total_male_10_to_14_years=sum(total_male_10_to_14_years),
          total_male_15_to_17_years=sum(total_male_15_to_17_years),
          total_male_18_and_19_years=sum(total_male_18_and_19_years),
          total_male_20_years=sum(total_male_20_years),
          total_male_21_years=sum(total_male_21_years),
          total_male_22_to_24_years=sum(total_male_22_to_24_years),
          total_male_25_to_29_years=sum(total_male_25_to_29_years),
          total_male_30_to_34_years=sum(total_male_30_to_34_years),
          total_male_35_to_39_years=sum(total_male_35_to_39_years),
          total_male_40_to_44_years=sum(total_male_40_to_44_years),
          total_male_45_to_49_years=sum(total_male_45_to_49_years),
          total_male_50_to_54_years=sum(total_male_50_to_54_years),
          total_male_55_to_59_years=sum(total_male_55_to_59_years),
          total_male_60_and_61_years=sum(total_male_60_and_61_years),
          total_male_62_to_64_years=sum(total_male_62_to_64_years),
          total_male_65_and_66_years=sum(total_male_65_and_66_years),
          total_male_67_to_69_years=sum(total_male_67_to_69_years),
          total_male_70_to_74_years=sum(total_male_70_to_74_years),
          total_male_75_to_79_years=sum(total_male_75_to_79_years),
          total_male_80_to_84_years=sum(total_male_80_to_84_years),
          total_male_85_years_and_over=sum(total_male_85_years_and_over),
          total_female=sum(total_female),
          total_female_under_5_years=sum(total_female_under_5_years),
          total_female_5_to_9_years=sum(total_female_5_to_9_years),
          total_female_10_to_14_years=sum(total_female_10_to_14_years),
          total_female_15_to_17_years=sum(total_female_15_to_17_years),
          total_female_18_and_19_years=sum(total_female_18_and_19_years),
          total_female_20_years=sum(total_female_20_years),
          total_female_21_years=sum(total_female_21_years),
          total_female_22_to_24_years=sum(total_female_22_to_24_years),
          total_female_25_to_29_years=sum(total_female_25_to_29_years),
          total_female_30_to_34_years=sum(total_female_30_to_34_years),
          total_female_35_to_39_years=sum(total_female_35_to_39_years),
          total_female_40_to_44_years=sum(total_female_40_to_44_years),
          total_female_45_to_49_years=sum(total_female_45_to_49_years),
          total_female_50_to_54_years=sum(total_female_50_to_54_years),
          total_female_55_to_59_years=sum(total_female_55_to_59_years),
          total_female_60_and_61_years=sum(total_female_60_and_61_years),
          total_female_62_to_64_years=sum(total_female_62_to_64_years),
          total_female_65_and_66_years=sum(total_female_65_and_66_years),
          total_female_67_to_69_years=sum(total_female_67_to_69_years),
          total_female_70_to_74_years=sum(total_female_70_to_74_years),
          total_female_75_to_79_years=sum(total_female_75_to_79_years),
          total_female_80_to_84_years=sum(total_female_80_to_84_years),
          total_female_85_years_and_over=sum(total_female_85_years_and_over)) %>%
mutate(total_male_change_ave=total_male_change_sum*100/total_male,
       total_male_under_5_years_change_ave=total_male_under_5_years_change_sum*100/total_male_under_5_years,
       total_male_5_to_9_years_change_ave=total_male_5_to_9_years_change_sum*100/total_male_5_to_9_years,
       total_male_10_to_14_years_change_ave=total_male_10_to_14_years_change_sum*100/total_male_10_to_14_years,
       total_male_15_to_17_years_change_ave=total_male_15_to_17_years_change_sum*100/total_male_15_to_17_years,
       total_male_18_and_19_years_change_ave=total_male_18_and_19_years_change_sum*100/total_male_18_and_19_years,
       total_male_20_years_change_ave=total_male_20_years_change_sum*100/total_male_20_years,
       total_male_21_years_change_ave=total_male_21_years_change_sum*100/total_male_21_years,
       total_male_22_to_24_years_change_ave=total_male_22_to_24_years_change_sum*100/total_male_22_to_24_years,
       total_male_25_to_29_years_change_ave=total_male_25_to_29_years_change_sum*100/total_male_25_to_29_years,
       total_male_30_to_34_years_change_ave=total_male_30_to_34_years_change_sum*100/total_male_30_to_34_years,
       total_male_35_to_39_years_change_ave=total_male_35_to_39_years_change_sum*100/total_male_35_to_39_years,
       total_male_40_to_44_years_change_ave=total_male_40_to_44_years_change_sum*100/total_male_40_to_44_years,
       total_male_45_to_49_years_change_ave=total_male_45_to_49_years_change_sum*100/total_male_45_to_49_years,
       total_male_50_to_54_years_change_ave=total_male_50_to_54_years_change_sum*100/total_male_50_to_54_years,
       total_male_55_to_59_years_change_ave=total_male_55_to_59_years_change_sum*100/total_male_55_to_59_years,
       total_male_60_and_61_years_change_ave=total_male_60_and_61_years_change_sum*100/total_male_60_and_61_years,
       total_male_62_to_64_years_change_ave=total_male_62_to_64_years_change_sum*100/total_male_62_to_64_years,
       total_male_65_and_66_years_change_ave=total_male_65_and_66_years_change_sum*100/total_male_65_and_66_years,
       total_male_67_to_69_years_change_ave=total_male_67_to_69_years_change_sum*100/total_male_67_to_69_years,
       total_male_70_to_74_years_change_ave=total_male_70_to_74_years_change_sum*100/total_male_70_to_74_years,
       total_male_75_to_79_years_change_ave=total_male_75_to_79_years_change_sum*100/total_male_75_to_79_years,
       total_male_80_to_84_years_change_ave=total_male_80_to_84_years_change_sum*100/total_male_80_to_84_years,
       total_male_85_years_and_over_change_ave=total_male_85_years_and_over_change_sum*100/total_male_85_years_and_over,
       total_female_change_ave=total_female_change_sum*100/total_female,
       total_female_under_5_years_change_ave=total_female_under_5_years_change_sum*100/total_female_under_5_years,
       total_female_5_to_9_years_change_ave=total_female_5_to_9_years_change_sum*100/total_female_5_to_9_years,
       total_female_10_to_14_years_change_ave=total_female_10_to_14_years_change_sum*100/total_female_10_to_14_years,
       total_female_15_to_17_years_change_ave=total_female_15_to_17_years_change_sum*100/total_female_15_to_17_years,
       total_female_18_and_19_years_change_ave=total_female_18_and_19_years_change_sum*100/total_female_18_and_19_years,
       total_female_20_years_change_ave=total_female_20_years_change_sum*100/total_female_20_years,
       total_female_21_years_change_ave=total_female_21_years_change_sum*100/total_female_21_years,
       total_female_22_to_24_years_change_ave=total_female_22_to_24_years_change_sum*100/total_female_22_to_24_years,
       total_female_25_to_29_years_change_ave=total_female_25_to_29_years_change_sum*100/total_female_25_to_29_years,
       total_female_30_to_34_years_change_ave=total_female_30_to_34_years_change_sum*100/total_female_30_to_34_years,
       total_female_35_to_39_years_change_ave=total_female_35_to_39_years_change_sum*100/total_female_35_to_39_years,
       total_female_40_to_44_years_change_ave=total_female_40_to_44_years_change_sum*100/total_female_40_to_44_years,
       total_female_45_to_49_years_change_ave=total_female_45_to_49_years_change_sum*100/total_female_45_to_49_years,
       total_female_50_to_54_years_change_ave=total_female_50_to_54_years_change_sum*100/total_female_50_to_54_years,
       total_female_55_to_59_years_change_ave=total_female_55_to_59_years_change_sum*100/total_female_55_to_59_years,
       total_female_60_and_61_years_change_ave=total_female_60_and_61_years_change_sum*100/total_female_60_and_61_years,
       total_female_62_to_64_years_change_ave=total_female_62_to_64_years_change_sum*100/total_female_62_to_64_years,
       total_female_65_and_66_years_change_ave=total_female_65_and_66_years_change_sum*100/total_female_65_and_66_years,
       total_female_67_to_69_years_change_ave=total_female_67_to_69_years_change_sum*100/total_female_67_to_69_years,
       total_female_70_to_74_years_change_ave=total_female_70_to_74_years_change_sum*100/total_female_70_to_74_years,
       total_female_75_to_79_years_change_ave=total_female_75_to_79_years_change_sum*100/total_female_75_to_79_years,
       total_female_80_to_84_years_change_ave=total_female_80_to_84_years_change_sum*100/total_female_80_to_84_years,
       total_female_85_years_and_over_change_ave=total_female_85_years_and_over_change_sum*100/total_female_85_years_and_over)

wirte.csv(reddf, "age_group_changes.csv")
```

This section plots the results from the previous code chunk.

```{r, fig.width = 15, fig.height = 4}
library(ggplot2)
library(tidyverse)

age_groups <- read.csv("age_group_changes.csv")
age_groups_fixed <- pivot_longer(age_groups, cols = 2:3, names_to = "gender", values_to = "coef")
ggplot(data = age_groups_fixed, aes(x = X, y = coef, fill = gender)) +
  geom_bar(position = "dodge", stat = "identity") +
  scale_x_discrete(limits=age_groups_fixed$X) +
  labs(y= "Reduction of PM2.5 \n (in Percent)", x = "Age Group") +
  scale_fill_manual(values=c("#b84279","#3751e5"))
```

## Conclusion

Each age group, regardless of gender, experienced an average reduction in PM2.5 of about 16-17%. Younger age groups showed a slightly higher reduction, potentially due to greater reliance on public transportation and their closer proximity to the light rail. This disparity suggests a need for further research into the lesser reductions observed in older age groups.

## References

Similar Paper (Used for some of the methodology)

1.  Chen, Y., & Whalley, A. (2012). Green infrastructure: The effects of urban rail transit on air quality. *American Economic Journal: Economic Policy*, *4*(1), 58-97.

Sources of Data

1.  Li, B., H. Beaudoing, and M. Rodell, NASA/GSFC/HSL (2020), GLDAS Catchment Land Surface Model L4 daily 0.25 x 0.25 degree GRACE-DA1 V2.2, Greenbelt, Maryland, USA, Goddard Earth Sciences Data and Information Services Center (GES DISC), Accessed: 06/17/2024, [10.5067/TXBMLX370XX8](https://doi.org/10.5067/TXBMLX370XX8)

2.  Di, Q., Y. Wei, A. Shtein, C. Hultquist, X. Xing, H. Amini, L. Shi, I. Kloog, R. Silvern, J. Kelly, M. B. Sabath, C. Choirat, P. Koutrakis, A. Lyapustin, Y. Wang, L. J. Mickley, and J. Schwartz.

    2021. Daily and Annual PM2.5 Concentrations for the Contiguous United States, 1-km Grids, v1 (2000 - 2016). Palisades, New York: NASA Socioeconomic Data and Applications Center (SEDAC). <https://doi.org/10.7927/0rvr-4538>. Accessed 06/17/2024.

3.  Di, Q., H. Amini, L. Shi, I. Kloog, R. Silvern, J. Kelly, M. B. Sabath, C. Choirat, P. Koutrakis, A. Lyapustin, Y. Wang, L. J. Mickley, and J. Schwartz. 2019. An Ensemble-based Model of PM2.5 Concentration Across the Contiguous United States with High Spatiotemporal Resolution. Environment International 130: 104909. <https://doi.org/10.1016/j.envint.2019.104909>.

Sources for Relevant Policies:

1.  Minnesota Public Utilities Commission. (2007). Renewable Energy Standard. Retrieved from [https://mn.gov/puc/activities/economic-analysis/renewable-environmental-dockets/#:\~:text=Renewable%20energy%20goals%20were%20passed,%2C%20the%20state's%20largest%20utility](https://mn.gov/puc/activities/economic-analysis/renewable-environmental-dockets/#:~:text=Renewable%20energy%20goals%20were%20passed,%2C%20the%20state's%20largest%20utility)

2.  Vehicle Emissions Inspection and Maintenance (I/M): Policy and technical guidance \| US EPA. (2024, February 2). US EPA. <https://www.epa.gov/state-and-local-transportation/vehicle-emissions-inspection-and-maintenance-im-policy-and-technical>

3.  Final Rule for Control of Air Pollution from New Motor Vehicles: Heavy-Duty Engine and Vehicle Standards and Highway Diesel Fuel Sulfur Control Requirements \| US EPA. (2024, May 17). US EPA. <https://www.epa.gov/regulations-emissions-vehicles-and-engines/final-rule-control-air-pollution-new-motor-vehicles>

4.  Proposed Rule and Correction for control of Air pollution from new motor vehicles: In-Use Testing for Heavy-Duty Diesel Engines and Vehicles \| US EPA. (2023, July 11). US EPA. <https://www.epa.gov/regulations-emissions-vehicles-and-engines/final-rule-control-air-pollution-new-motor-vehicles-heavy>

5.  Final rule for control of emissions of air pollution from nonroad diesel engines and fuel \| US EPA. (2024, February 6). US EPA. <https://www.epa.gov/regulations-emissions-vehicles-and-engines/final-rule-control-emissions-air-pollution-nonroad#:~:text=Rule%20Summary%20In%202004%2C%20EPA%20finalized%20Tier,nonattainment%20areas%20to%20improve%20their%20air%20quality>.

6.  Reducing mercury releases \| Minnesota Pollution Control Agency. (n.d.). Minnesota Pollution Control Agency. <https://www.pca.state.mn.us/air-water-land-climate/reducing-mercury-releases>

7.  Clean Cars Minnesota. (n.d.). Conservation MN. <https://www.conservationminnesota.org/clean-cars-minnesota>

8.  2030 Regional Development Framework - Metropolitan Council. (n.d.). Metropolitan Council. <https://metrocouncil.org/Planning/Thrive-2040/2030-Regional-Development-Framework.aspx>

9.  Environmental Initiative. (2024, April 19). Clean Air Minnesota - Environmental Initiative. <https://environmental-initiative.org/our-work/clean-air-minnesota/>

Sources for Confounding Sources of Pollution:

1.  Minnesota Pollution Control Agency. (n.d.). Northern Metal Recycling: Air Quality Permit Violations. Retrieved 2024, from <https://www.pca.state.mn.us/news-and-stories/northern-metals-pays-12000-for-violating-air-quality-regulations-at-its-becker-facility#:~:text=state.mn.us-,Northern%20Metals%20pays%20%2412%2C000%20for%20violating%20air%20quality%20regulations%20at,particulate%20matter%20>(PM10%20and%20PM2.

2.  St Anthony, N., & Tribune, S. (2022, February 13). Flint Hills’ Pine Bend refinery produces more fuel with less pollution. Star Tribune. <https://www.startribune.com/flint-hills-pine-bend-refinery-produces-more-fuel-with-less-pollution/600146396/>

3.  Xcel Energy. (n.d.). <https://mn.my.xcelenergy.com/s/about/newsroom/press-release/xcel-energy-takes-major-step-toward-replacing-coal-fired-king-plant-with-clean-e-MCQ6MJVC5VIZCSLMVOCUPTC5KMQM>

4.  AIR EMISSION PERMIT. (2008). In epa.gov (No. 12300055–004). Environmental Protection Agency. <https://www.epa.gov/system/files/documents/2021-12/gerdauameristeel_12300055-004.pdf>

5.  Minnesota Department of Health. (2004). Public health assessment for perfluorochemical contamination at a hazardous waste site in Minnesota. In Minnesota Department of Health Report. <https://www.health.state.mn.us/communities/environment/hazardous/docs/sites/washington/3mcg0205.pdf>

6.  Hennepin County wrestles with future of Minneapolis waste incinerator. (2023, September 21). MPR News. <https://www.mprnews.org/story/2023/09/21/hennepin-county-wrestles-with-future-of-minneapolis-waste-incinerator>

7.  APA Citation: KOCH REFINING CO./N-REN CORP. \| Superfund Site Profile \| Superfund Site Information \| US EPA. (n.d.). <https://cumulis.epa.gov/supercpad/SiteProfiles/index.cfm?fuseaction=second.cleanup&id=0503686>

8.  APA Citation: Hillesheim, D. V. & Xcel Energy. (2023). Revised Petition for Alternative Operational Limits for Continuous Compliance Demonstration under 40 C.F.R. Part 63, Subpart YYYY. In U.S. Environmental Protection Agency \[Report\]. <https://www.epa.gov/system/files/documents/2023-09/Xcel%20Energy%20High%20Bridge_05-22-2023.pdf>

9.  Rademacher, N. (n.d.). Meet the Main Energy Plant: a hidden world reducing University emissions. The Minnesota Daily. <https://mndaily.com/257087/news/acenergyplant/>

10. ANOKa \| Federal Ammunition \| Minnesota Pollution Control Agency. (n.d.). Minnesota Pollution Control Agency. <https://www.pca.state.mn.us/local-sites-and-projects/anoka-federal-ammunition>

11. Sfiecke. (2006, September 8). EPA cites CertainTeed for clean air violations. SWNewsMedia.com. <https://www.swnewsmedia.com/shakopee_valley_news/news/epa-cites-certainteed-for-clean-air-violations/article_ad8d6f3b-44f5-59ab-8451-c3860538a7a9.html>

12. EJMap. (n.d.). <https://ejmap.org/displayfacility-74250.htm?giFacilityid=74250&gsTable=facility>
